---
title: "confidence interval for bootstrap"
author: "Parnian Jahangiri Rad"
date: "1/7/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
set.seed(9813007)
n <- 16
mu <- 7
sd <- 2
X <- rnorm(n,mu,sd)
theta.hat <- mean(X)
theta.hat
```

## Confidence interval for **mean**:  
```{r}
#Bootstrap
nboot <- 1000
theta.hat.star <- numeric(nboot)
for(i in 1:nboot){
  y <- sample(X,n,replace = T)
  theta.hat.star[i] <- mean(y)
}
theta.hat.star <- sort(theta.hat.star)
hist(theta.hat.star)
```
95% confidence interval for $\mu$ when we have standard error:  
```{r}
#confidence interval with 95% confidence when we do have se
x.bar <- mean(X)
a <- (x.bar - (1.96 * (sd/sqrt(n))))
b <- (x.bar + (1.96 * (sd/sqrt(n))))
cbind(a,b)
abs(a-b)
```

```{r}
#confidence interval with 95% confidence using quantile(for X sample)
#This method is exact
d = quantile(theta.hat.star, c(0.025, 0.975)) 
d
```
### The Percentile Bootstrap method  
$\alpha=0.05$ for 95% CIs:  
```{r}
alpha <- 0.05
j <- (alpha/2)*nboot
k <- (1- alpha/2)*nboot
cbind(theta.hat.star[j],theta.hat.star[k])
#interval's length
interval.length1 <- theta.hat.star[k] - theta.hat.star[j]
interval.length1
```
### The Bootstrap t interval method  
```{r}
theta.hats.b <- numeric(nboot)
ts <- numeric(nboot)

for(b in 1:nboot){
  y <- sample(X, size = n, replace = TRUE)
  theta.hats.b[b] <- mean(y)
  sd.theta.hats.b <- numeric(nboot)
  
  for(b2 in 1:nboot){
    y2 <- sample(X,size = n,replace = TRUE)
    sd.theta.hats.b[b2] <- mean(y2)
  }
  
  se.b <- sd(sd.theta.hats.b)
  
   ts[b] <- (theta.hats.b[b] - theta.hat) / se.b
}

alpha <- 0.05
ts.ordered <- sort(ts)

qs <- quantile(ts.ordered, probs = c(alpha/2, 1-(alpha/2)))
se.hat <- sd(theta.hats.b)

(CI = c(theta.hat - qs[2]*se.hat, theta.hat - qs[1]*se.hat))
#interval's length
interval.length2 <- CI[2] - CI[1]
```

```{r}
interval.length1
interval.length2
```
**Bootstrap t interval** method is more accurate than percentile since
it has a smaller interval length.(and confidence interval for both
methods is 95%.)  

## Confidence interval for **variance**:  
Now, our $\hat{\theta}$ is variance:  
```{r}
theta.hat <- var(X)
theta.hat
```

```{r}
#Bootstrap
nboot <- 1000
theta.hat.star <- numeric(nboot)
for(i in 1:nboot){
  y <- sample(X,n,replace = T)
  theta.hat.star[i] <- var(y)
}
theta.hat.star <- sort(theta.hat.star)
hist(theta.hat.star)
```

```{r}
#confidence interval with 95% confidence using quantile(for X sample)
#This method is exact
d = quantile(theta.hat.star, c(0.025, 0.975)) 
d
```

### The Percentile Bootstrap method   
$\alpha=0.05$ for 95% CIs:  
```{r}
alpha <- 0.05
j <- (alpha/2)*nboot
k <- (1- alpha/2)*nboot
cbind(theta.hat.star[j],theta.hat.star[k])
#interval's length
interval.length1 <- theta.hat.star[k] - theta.hat.star[j]
```
### ### The Bootstrap t interval method  
```{r}
theta.hats.b <- numeric(nboot)
ts <- numeric(nboot)

for(b in 1:nboot){
  y <- sample(X, size = n, replace = TRUE)
  theta.hats.b[b] <- var(y)
  sd.theta.hats.b <- numeric(nboot)
  
  for(b2 in 1:nboot){
    y2 <- sample(X,size = n,replace = TRUE)
    sd.theta.hats.b[b2] <- var(y2)
  }
  
  se.b <- sd(sd.theta.hats.b)
  
   ts[b] <- (theta.hats.b[b] - theta.hat) / se.b
}

alpha <- 0.05
ts.ordered <- sort(ts)

qs <- quantile(ts.ordered, probs = c(alpha/2, 1-(alpha/2)))
se.hat <- sd(theta.hats.b)

(CI = c(theta.hat - qs[2]*se.hat, theta.hat - qs[1]*se.hat))
#interval's length
interval.length2 <- CI[2] - CI[1]
```

```{r}
interval.length1
interval.length2
```
Again, **Bootstrap t interval** method is more accurate than percentile since it has a smaller interval length.(and confidence interval for both
methods is 95%.)  

